---
title: "Available Hooks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{available-hooks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


Below is a comprehensive list with all hooks from {precommit} as well well as
their arguments or flags, if they take any. Other repos also host hooks, many
are listed [here](https://pre-commit.com/hooks.html).

# Good to know

**File modification**

Some hooks will fail without changing the files you want to commit, like the
`lintr` hook - and you need to make manual changes for the hook to pass on the
next attempt. Other hooks like the `styler` hook write to files, and if that
changes the file, the hook will fail, but this means you won't need to modify
the file manually, just stage the changes and try to commit again. Below, we
indicate for every hook if it modifies files or not.

**Arguments**

Arguments are specified as described in the [pre-commit.com
documentation](https://pre-commit.com/#passing-arguments-to-hooks), e.g. we can
set the argument `style_pkg` to `styler` and `style_fun` to `tidyverse_style`
for the hook `style-files` like this:

```{r, echo = FALSE, output = "asis", comment = "", message = FALSE}
library(magrittr)
rev <- paste0("v", packageVersion("precommit"))
cat(glue::glue("
repos:
-   repo: https://github.com/lorenzwalthert/precommit
    rev: {rev}
    hooks:
    -   id: style-files
        args: [--style_pkg=styler, --style_fun=tidyverse_style]
"))
```

**Other hook settings**

Apart from specifying the `args` key as described above, there are other hooks
settings you can specify. If you don't, they are inherited from the default 
repository's configuration (i.e. the `.pre-commit-hooks.yaml` file in 
https://github.com/lorenzwalthert/precommit). See the [pre-commit documentation](https://pre-commit.com/#pre-commit-configyaml---hooks), for the
available settings.

# Hooks

## `style-files`

A hook to style files with [styler](https://styler.r-lib.org). Only commit code
corresponding to the tidyverse style guide. Set argument `style_pkg` and
`style_fun` if you want to use another style guide than the tidyverse style
guide.

```
  id: style-files
  args: [--style_pkg=<styler>, --style_fun=<tidyverse_style>]
```

This hook modifies files.

## `readme-rmd-rendered`

Make sure `README.Rmd` hasn't been edited more recently than `README.md`, i.e.
remind you to render the `.Rmd` to `.md` before committing.

This hook does not modify files.

## `parsable-R`

Checks if your `.R` files are "valid" R code by checking if running `parse()` on
it returns an error.

This hook does not modify files.

## `no-browser-statement`

Guarantees you that you don't accidentally commit code with a `browser()`
statement in it.

This hook does not modify files.

## `spell-check`

Checks spelling with `spelling::spell_check_files()`. 

**Excluded files**

When you invoke `precommit::use_precommit()` and `.pre-commit-config.yaml` is
written to your repo (unless you specify `config_source` otherwise), we copy 
the expression in the `exclude:` key from
spell check hook the default repository's configuration (i.e. the 
`.pre-commit-hooks.yaml` file in https://github.com/lorenzwalthert/precommit)
into your config file, so you can easily add or remove some files. As of 
`r paste0("v", packageVersion("precommit"))`, the following regex is used to 
exclude files following the 
[verbose python regex syntax](https://pre-commit.com/#regular-expressions):

```{r, echo = FALSE, comment = ""}
readLines(system.file("pre-commit-hooks.yaml", package = "precommit")) %>%
  gsub("^ *exclude *: *>", "    exclude: |", .) %>%
  yaml::yaml.load() %>%
  purrr::keep(~ .x$id == "spell-check") %>%
  magrittr::extract2(1) %>%
  magrittr::extract2("exclude") %>%
  cat(sep = "\n")
```

**language**

The `lang` arg will be passed to `spelling::spell_check_files()`.

```
id: spell-check
args: [--lang=<language>]
exclude: <regular expresion>
```

This hook does not modify input files. It will add all words not found in the
dictionary to `inst/WORDLIST`, assuming they were spelled correctly but were not
in the dictionary. An example might be "RStudio". The hook error message will
contain all words written to `inst/WORDLIST`, so if there were really some
typos, make sure to fix them and remove them from `inst/WORDLIST`. If there were
not typos, or you fixed all, stage `inst/WORDLIST` and this time, the commit
should pass.

## `roxygenize`

A hook to run `roxygen2::roxygenize()`. Makes sure you commit your `.Rd` changes
with the source changes.

This hook does not modify input files, but writes to `.Rd` files in `man/`.

## `deps-in-desc`

Checks if packages used with the `pkgname::fun()` syntax are listed in your
DESCRIPTION file. Flag `allow_private_imports` lets the user specify that
private imports into the package namespace are tolerable, e.g. `somepkg:::x()`.
Flag not set by default, i.e. the hook will fail if such a call is found.

```
id: deps-in-desc args: [--allow_private_imports] 
```

This hook does not modify the file `DESCRIPTION` because the user should decide
for each package if it should go to `Imports:` or `Suggests:`, which can be done
easily with `usethis::use_package()`. For those who use
`usethis::use_package()`: As long as the CRAN version of the dependency {desc}
is `v1.2.0`, we recommend you to install the GitHub version because otherwise,
the new dependencies won't be added in alphabetical order and after adding a
dependency, `use-tidy-description` will fail in most cases
(#85)[https://github.com/r-lib/desc/pull/85].

## `use-tidy-description`

A hook to run `usethis::use_tidy_description()` to ensure dependencies are
ordered alphabetically and fields are in standard order.

This hook does modify the file `DESCRIPTION`.

## `lintr`

A hook to run `lintr::lint()` to check that R files are lint free. Argument
`warning_only` changes the behavior of the pre-commit to be non-blocking. You
should set this with the field `verbose: true`.

```
id: lintr
args: [--warn_only]
verbose: true
```

When configured this way, lintr prints lint errors as they appear. Other
arguments are not supported. Instead, `lintr` config should be specified in a
`.lintr` config file in Debian Control Field Format as specified in the
[`.lintr` documentation](https://github.com/jimhester/lintr#project-configuration).

This hook does not modify any file.

## `codemeta-description-updated`

Make sure `DESCRIPTION` hasn't been edited more recently than `codemeta.json`,
i.e. remind you to run `codemetar::write_codemeta()` in order to keep
`codemeta.json` in sync with `DESCRIPTION`.

This hook does not modify any file.
