---
title: "Available Hooks"
description: > 
  Check out all hooks this repo contains and how they can be customized
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{available-hooks}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

Below is a comprehensive list with all hooks from {precommit} as well as
their arguments or flags, if they take any. With the standard config
file that gets placed in your project with `precommit::use_precommit()`
all hooks should work out of the box, but you can further customize them
as described below. Other repos also host hooks, many are listed
[here](https://pre-commit.com/hooks.html).

# Good to know

**File modification**

Some hooks will fail without changing the files you want to commit, like
the `lintr` hook - and you need to make manual changes for the hook to
pass on the next attempt. Other hooks like the `roxygenize` hook write
to files, and if that changes the file, the hook will fail, but this
means for most hooks you won't need to modify the file manually after
the attempted commit, just stage the changes and try to commit again.
Below, we indicate for every hook if it modifies files or not.

**Arguments**

Arguments for the hooks are specified as described in the
[pre-commit.com
documentation](https://pre-commit.com/#passing-arguments-to-hooks).[^1]
You can specify arguments like this:

[^1]: Note that there might be issues with arguments that contain
    special characters and you might have to quote them and the order of
    single, double and escaped quotes may not give identical results on
    one platform and may not be portable to all platforms either.

```{r, echo = FALSE, output = "asis", comment = "", message = FALSE}
library(magrittr)
rev <- paste0("v", packageVersion("precommit"))
cat(glue::glue("
repos:
-   repo: https://github.com/lorenzwalthert/precommit
    rev: {rev}
    hooks:
    -   id: lintr
        args: [--warn-only, --key=value]
"))
```

**Other hook settings**

Apart from specifying the `args` key as described above, there are other
hooks settings you can specify, e.g. file exclusion. If you don't, they
are inherited from the default repository's configuration (i.e. the
`.pre-commit-hooks.yaml` file in
<https://github.com/lorenzwalthert/precommit>). See the [pre-commit
documentation](https://pre-commit.com/#pre-commit-configyaml---hooks),
for the available settings.

# Hooks

## `style-files`

A hook to style files with [styler](https://styler.r-lib.org). Only
commit code corresponding to the tidyverse style guide. You can pass
arguments to
[`style_file(...)`](https://styler.r-lib.org/reference/style_file.html)
using the `--key=value` syntax like this:

      id: style-files
      args: [--scope=spaces, --reindention=specify_reindention('#')]

In addition, the hook takes the following arguments that are not passed
to
[`style_file(...)`](https://styler.r-lib.org/reference/style_file.html):

-   Argument `style_pkg` and `style_fun` default to `styler` and
    `tidyverse_style`. If you want to use another style guide than the
    tidyverse style guide, you can specify it like this:

<!-- -->

      id: style-files
      args: [--style_pkg=pkg.with.style.guide, --style_fun=exported.style.function]

-   The argument `--no-warn-cache` is deprecated and will be removed in
    a future release. Please remove it from your
    `.pre-commit-config.yaml`.

-   Argument `cache-root` is passed to `options()` to set
    `styler.cache_root`. Default value: `styler-perm`. The argument
    determines the sub-directory under the {R.cache} cache directory
    that {styler} uses. If you want {styler} to auto-clean up caches
    older than 6 days, set this to `"styler"`. For more information, see
    `help("caching", package = "styler")`.

<!-- -->

      id: style-files
      args: [--cache-root=styler]

-   Argument `ignore-start` and `ignore-stop` is passed to `options()`
    to set `styler.ignore_start` and `styler.ignore_stop` respectively.
    Not set by default, so the {styler} defaults apply. This was
    introduced in {precommit} 0.2.2.9012. For example, if you want to
    restore old behavior of styler \< 1.6.2, where only the literals
    `styler: off` and `styler: on` were accepted, use this regex:

<!-- -->

      id: style-files
      args: [--ignore-start="^# styler: on$", --ignore-stop="^# styler: off$"]

This hook modifies files unless you specify `--dry=fail` (requires
`{styler} > 1.3.2`).

## `readme-rmd-rendered`

Make sure `README.Rmd` hasn't been edited more recently than
`README.md`, i.e. remind you to render the `.Rmd` to `.md` before
committing.

This hook does not modify files.

## `parsable-R`

Checks if your `.R` and `.Rmd` files are "valid" R code by checking if
running `parse()` on them (or their `knitr::purl()`ed output for `.Rmd`)
returns an error.

This hook does not modify files.

## `no-browser-statement`

Guarantees you that you don't accidentally commit code with a
`browser()` statement in it.

This hook does not modify files.

## `no-debug-statement`

Guarantees you that you don't accidentally commit code with a `debug()`
or `debugonce()` statement in it.

This hook does not modify files. This hook was added in version
0.2.2.9012.

## `spell-check`

Checks spelling with `spelling::spell_check_files()`.

**Excluded files**

When you invoke `precommit::use_precommit()` and
`.pre-commit-config.yaml` is written to your repo (unless you specify
`config_source` otherwise), we copy the expression in the `exclude:` key
from spell check hook the default repository's configuration (i.e. the
`.pre-commit-hooks.yaml` file in
<https://github.com/lorenzwalthert/precommit>) into your config file, so
you can easily add or remove some files. As of
`r paste0("v", packageVersion("precommit"))`, the following regex is
used to exclude files following the [verbose python regex
syntax](https://pre-commit.com/#regular-expressions):

```{r, echo = FALSE, comment = ""}
readLines(system.file("pre-commit-hooks.yaml", package = "precommit")) %>%
  gsub("^ *exclude *: *>", "    exclude: |", .) %>%
  yaml::yaml.load() %>%
  purrr::keep(~ .x$id == "spell-check") %>%
  magrittr::extract2(1) %>%
  magrittr::extract2("exclude") %>%
  cat(sep = "\n")
```

**language**

The `lang` arg will be passed to `spelling::spell_check_files()`.

      id: spell-check
      args: [--lang=<language>]

This hook does not modify input files. It will add all words not found
in the dictionary to `inst/WORDLIST`, assuming they were spelled
correctly but were not in the dictionary. An example might be "RStudio".
The hook error message will contain all words written to
`inst/WORDLIST`, so if there were really some typos, make sure to fix
them and remove them from `inst/WORDLIST`. If there were not typos, or
you fixed all, stage `inst/WORDLIST` and this time, the commit should
pass.

## `roxygenize`

A hook to run `roxygen2::roxygenize()`. Makes sure you commit your `.Rd`
changes with the source changes. To take advantage of caching, you don't
need to run `roxygen2::roxygenize()` manually anymore. The argument
`--no-warn-cache` is deprecated and will be removed in a future release.
Please remove it from your `.pre-commit-config.yaml`.

Because the hook will write the version of {roxygen2} to `DESCRIPTON`,
you should either make sure the version you use when you call {roxygen2}
interactively matches the one from in {precommit} or simply not run
{roxygen2} manually.

If you specify additional roclets through the `Roxygen:` field in
`DESCRIPTION`, e.g. from [{pkgapi}](https://github.com/r-lib/pkgapi) you
must specify the dependencies explicitly such that `renv::install()`
understands it, e.g.

      id: roxygenize
      additional_dependencies:
      - r-lib/pkgapi

This hook does not modify input files, but writes to `.Rd` files in
`man/`, `NAMESPACE` and potentially others depending on which roxygen
roclets you specified in `DESCRIPTION`.

## `deps-in-desc`

Checks if packages used with the `pkgname::fun()` syntax are listed in
your DESCRIPTION file. Note that `README.Rmd` is never checked. Flag
`allow_private_imports` lets the user specify that private imports into
the package namespace are tolerable, e.g. `somepkg:::x()`. Flag not set
by default, i.e. the hook will fail if such a call is found.

      id: deps-in-desc 
      args: [--allow_private_imports] 

This hook does not modify the file `DESCRIPTION` because the user should
decide for each package if it should go to `Imports:` or `Suggests:`,
which can be done easily with `usethis::use_package()`.

## `use-tidy-description`

A hook to run `usethis::use_tidy_description()` to ensure dependencies
are ordered alphabetically and fields are in standard order.

This hook does modify the file `DESCRIPTION`.

## `lintr`

A hook to run `lintr::lint()` to check that R files are lint free.
Argument `warning_only` changes the behavior of the pre-commit to be
non-blocking. You should set this with the field `verbose: true`.

      id: lintr
      args: [--warn_only]
      verbose: true

When configured this way, lintr prints lint errors as they appear. Other
arguments are not supported. Instead, `lintr` config should be specified
in a `.lintr` config file in Debian Control Field Format as specified in
the [`.lintr`
documentation](https://github.com/r-lib/lintr#project-configuration).

This hook does not modify any file.

## `codemeta-description-updated`

Make sure `DESCRIPTION` hasn't been edited more recently than
`codemeta.json`,

i.e. remind you to run `codemetar::write_codemeta()` in order to keep
`codemeta.json` in sync with `DESCRIPTION`.

This hook does not modify any file.
