---
title: "Data Structures"
author: "Lorenz Walthert and Kirill Müller"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette illustrates how the core of `styler` currently^[at commit e`6ddee0f510d3c9e3e22ef68586068fa5c6bc140`] works, i.e. how
rules are applied to a parse table and how limitations of this approach can be 
overcome with a refined approach.

## Status quo - the flat approach

Roughly speaking, a string containing code to be formatted is parsed with `parse`
and the output is passed to `getParseData` in order to obtain a parse
table with detailed information about every token. For a simple example string
"`a <- function(x) { 1+1 }`" to be formatted, the parse table on which `styler`
performs the manipulations looks similar to the one presented below.
```{r, message = FALSE}
library("styler")
library("dplyr")

code <- "a <- function(x) { if(x > 1) { 1+1 } else {x} }"

(parse_table <- compute_parse_data_flat_with_ws(code))
```
The column `spaces` was computed from the columns `col1` and `col2`, `newlines`
was computed from `line1` and `line2` respectively.

So far, styler can set the spaces around the operators correctly. In our example, 
that involves adding spaces around `+`, so in the `spaces` column, element nine
and ten must be set to one. This means that a space is added after `1` and after `+`. 
To get the spacing right and cover the various cases, a set of functions has to 
be applied to the parse table subsequently (and in the right order), 
which is essentially done via `Reduce()`. 
After all modifications on the table are completed, `serialize_parse_data()`
collapses the `text` column and adds the number of spaces and 
line breaks specified in `spaces` and `newlines` in between the elements of
`text`. If we serialize our table and don't perform any modification, we 
obviously just get back what we started with.
```{r}
serialize_parse_data(parse_table)
```

## Refining the flat approach - nesting the parse table

Although the flat approach is good place to start, e.g. for fixing spaces
between operators, it has its limitations. In particular, it treats each token 
the same way in the sense that it does not account for the context of the token, 
i.e. in which sub-expression it appears.
To set the indention correctly, we need a hierarchical view on the parse data, 
since all tokens in a sub-expression have the same indention level. Hence, 
a natural approach would be to create a nested parse table instead of a flat
parse table and than take a recursion over all elements in the table, so for 
each sub(-sub etc.)-expression, a separate parse table would be created and the 
modifications would be applied to this table before putting everything back 
together. A function to create a nested parse table already exists in `styler`.
Let's have a look at the top level:

```{r}
(l1 <- compute_parse_data_nested(code)[-1])

```

The tibble contains the column `child`, which itself contains a tibble. 
If we "enter"  the first child, we can see that the expression was split up 
further.

```{r}
l1$child[[1]] %>%
  select(text, terminal, child, token)
```

And further...
```{r}
l1$child[[1]]$child[[3]]$child[[5]]$child[[2]]
```

... and so on. Every child that is not a terminal contains another tibble where 
the sub-expression is split up further - until we are left with tibbles that 
only contain terminals.


Recall the above example. `1+1` 
is a sub-expression of the expression `code`, and we would like to set the 
indention for `1+1`, even before we do anything with `1+1`.

The code for a function that kind of sketches the idea and illustrates such a 
recursion is given below: 

```{r}
serialize <- function(x) {
  out <- Map(
    function(terminal, token, child) {
      if (terminal)
        token
      else
        serialize(child)
    },
    x$terminal, x$token,
    if (is.null(x$child)) rep(list(NULL), length(x$terminal)) else x$child
  )
  out
}

x <- compute_parse_data_nested(code)
serialize(x) %>% unlist
```

How to exactly implement a similar recursion to not just return the tokens, but 
the styled code is subject to future work, so would be the functions to be
applied to a sub-expression parse table that create correct indention.

## Final Remarks

Although a flat structure would possibly also allow to solve the problem, it is
a less elegant and flexible solution to the problem, so the authors of this 
vignette think its worth giving the nested approach a try.
