---
title: "Alignment detection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Alignment detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Sometimes, you deliberately align code to make it more readable.

```{r}
call(
  a =       3,
  b = 3213232
)

```

Until styler 1.1.1.9002 (with `strict = TRUE`, e.g. as in
`styler_style_file(..., strict = TRUE)`), this was formatted as follows:

```{r}
call(
  a = 3,
  b = 3213232
)

```

because no alignment detection was built in. With `strict = FALSE`, the spacing
would have been kept, however, `strict = FALSE` has a number of other
implications because it is in general less invasive (e.g. it would not add
braces to the following expression, whereas `strict = TRUE` would):

```{r}
if (TRUE)
  call(another, arg)
```

Back to the initial topic, styler >= 1.1.1.9003 detects the aforementioned
alignment. This vignette describes how an aligned code is defined by styler and
gives some examples so users can format their aligned code to match the
definition styler uses to ensure their code is not unintentionally reformatted.

## function calls

### Non-technical

If all arguments in the first column are named: Make commas match position
vertically and right align everything between commas

```{r}
# all arguments of first column named -> must right align
# aligned if the (imaginary) comma on the last line is in line with the commas 
# from the two top lines.
fell(
  x  =    1, 
  y  =   23,
  zz = NULL
)
```

If not all arguments of the first column are named: Make all but the first
column's commas match position vertically and right align everything between the
commas, except before the first comma on a line, give priority to correctly
indent.

```{r}
# not all arguments of first column (f = NULL and 13231) named, hence, only 
# commas of all but the first column must agree.
gell(
  2,        g = gg(x), n  = 3 * 3,
  13231,  fds =    -1, gz = f / 3,
)
```

By *align everything in between the commas*, we mean put zero space before a
comma and at least one after. Note that the arguments on the first line are
ignored when detecting alignment, which is best shown when code is formatted
such that no line breaks will be modified by styler (which is the case if all
names on the first line are unnamed and all subsequent are named), like here:

```{r}
map(x, f, 
 arg1 =  121,
 arg2 =    1
)
```

## technical

Function calls are aligned if **all** of the following conditions hold (for all
but the very first line (i.e. `call(` below):

* first column has same number of lag spaces. This basically means that the
  indention is identical for all columns (except for the closing brace if it is
  on its own line). The below example has one column, because the maximal number
  of commas on one line is one.

```{r}
# holds
call(
  a =  3,
  b = 32
)

# doesn't hold
call(
  a =  3,
   b = 32
)
```

* spacing around comma (0 before, > 1 after, >= 0 after last column on line) and
  spacing around `=` (at least one before and after).

```{r}
# holds
call(
  a =  3, k  = 3,
  b = 32,    222
)

# doesn't hold
call(
  a =  3 ,
  b = 32
)
```

* All commas from all columns are aligned. This means that for every column, all
  commas must be on the same positions as the commas from the other lines. If
  not all arguments are named in the first column, this column is not considered.
  The reason to exclude the first column is that, as in the example below, it is
  possible that some arguments are named while others are not. Then, it is not
  generally possible to keep the first rule (i.e. indention identical across
  lines) as well as ensuring that the comma does not have any spaces before it
  and that the comma is aligned with the other lines. This is shown below with
  the line `f(x, y),`. For this reason, the requirements exclude the first
  column in such cases. The *holds* example shows that is is possible (but not
  required) for named arguments to also have the commas separating the first and
  second column aligned.

```{r}
# holds
call(
  a =  ff("pk"), k  = 3, x =  2,
  b =     f(-g), 22 + 1, yy = 1,
  c =         1,
  f(x, y), 
  k
)

# doesn't hold
call(
  a =   3,
  b = 32, c = 2
)
```

Note that the above definition does not check alignment of `=`, so styler will
treat the following as aligned:

```{r}
rge(
  x  = 99, x =  2,
  fs =  1,  y = 1,
)
```

## comments

not supported yet.

## assignment

not supported yet.

