---
title: "FAQ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FAQ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## How does it work? What's the magic?

No magic. Git [provides a mechanism for pre-commit
hooks](https://git-scm.com/book/en/v2/Customizing-Git-Git-Hooks). Since it's
cumbersome to copy bash scripts and keep them updated in some hidden `.git/`
directory, people wrote frameworks to basically maintain these files. One of
them is [pre-commit.com](https://pre-commit.com). So you call git to commit, git calls
its internal hooks, these hooks (bash scripts in the simplest case) are provided
by the pre-commit framework. And where does pre-commit get all it's nice hooks
from? From us. 

## How are hooks accessed? What happens if there's no internet connection?

Internet connection is required during installation 
(`precommit::install_precommit()`) and project initialization 
(`precommit::use_precommit()`) as well as when hooks are updated 
(`precommit::autoupdate()`). Otherwise, no internet connection is required. 
During initialization, hooks from this repo will be placed in a virtual 
environment within the user cache and upon committing, they are called from a 
generic hook helper in `.git/hooks/`.

## Why do some hooks not give me the same results as running the command interactively?

Note that starting with `precommit > 0.1.3`, hooks run in an isolated 
[`{renv}`](https://rstudio.github.io/renv/), leveraging [R as a supported language](https://pre-commit.com/#r) in the pre-commit framework. The version of a package you use interactively 
(e.g. `{roxygen2}` to document) might for that reason be different from the one 
pre-commit uses. The point of using pre-commit is to take mental burden off your
shoulders (and sometimes even caches results for speed), so there is no need to 
run functionality covered by the hooks manually anyways. If you need to know 
which package versions are used, look for the `renv.lock` file in 
https://github.com/lorenzwalthert/precommit at the git tag specified under 
`rev:` in your `.pre-commit-config.yaml`.

## Can you use it outside RStudio?

Yes, all but the `open_config()` and `open_wordlist()` to open files in RStudio. 

## Can I use the hooks provided in this package without installing the R package?

Yes, you don't need to manually install this package. Although technically 
speaking, this package will get installed into a virtual environment isolated 
from your global and project library.

## How can I make sure that my contributors are using the hooks?

They must follow the installation instructions in the README, i.e. run 
```{r}
remotes::install_github("lorenzwalthert/precommit")
precommit::install_precommit()
precommit::use_precommit()
```

The last call can be omitted by users who have [automatically enabled 
pre-commit hooks](https://pre-commit.com/#automatically-enabling-pre-commit-on-repositories).

To enforce all hooks pass, you can follow the [advice on how to use pre-commit in a CI/CD setup](https://pre-commit.com/#usage-in-continuous-integration).

## Should I list `{precommit}` as a dependency of my package in DESCRIPTION?

No, precommit is a dev dependency like `{devtools}`, so it should not be listed 
in `DESCRIPTION`.


## What if not all people who are committing to this repo want to use the hooks?

This is not a problem, git will only run the hooks in a local repo after
`precommit::use_precommit()` has been run successfully from within this local repo on your machine.
You can also uninstall anytime with `precommit::uninstall_precommit()`. Anyone
who does not want to use the hooks simply should not run
`precommit::use_precommit()`. You can also temporarily disable hooks using the 
environment variable `SKIP` or the argument `--no-verify`, as 
described [here](https://pre-commit.com/#temporarily-disabling-hooks).


## How does one create a new pre-commit hook?

How to contribute new hooks is explained in [`CONTRIBUTING.md`](https://github.com/lorenzwalthert/precommit/blob/main/CONTRIBUTING.md).

*** 

There is more. Check out the documentation of the [pre-commit](https://pre-commit.com) framework.
