---
title: "FAQ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FAQ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## How does it work? What's the magic?

No magic. Git [provides a mechanism for pre-commit
hooks](https://git-scm.com/book/en/v2/Customizing-Git-Git-Hooks). Since it's
cumbersome to copy bash scripts and keep them updated in some hidden `.git/`
directory, people wrote frameworks to basically maintain these files. One of
them is [pre-commit.com](https://pre-commit.com). So you call git to commit, git calls
its internal hooks, these hooks (bash scripts in the simplest case) are provided
by the pre-commit framework. And where does pre-commit get all it's nice hooks
from? From us.

## How are hooks accessed? What happens if there's no internet connection?

Internet connection is required during installation 
(`precommit::install_precommit()`) and project initialization 
(`precommit::use_precommit()`) as well as when hooks are updated 
(`precommit::autoupdate()`). Otherwise, no internet connection is required. 
Hooks will be placed into `.git/` during calling the 
`precommit::use_precommit()` and afterwards you can `precommit::autoupdate()` 
to replace the hooks with their new version (that's what I think happens).

## Can you use it outside RStudio?

Yes, all but the `open_config()` and `open_wordlist()` to open files in RStudio. 

## Can I use the hooks provided in this package without installing the R package?

Yes, apart from the `roxygenize` hook (which uses an exported function from 
{precommit}). The hooks work even if you don't have the R package installed 
because all you need to have for the hooks to work is the pre-commit executable. 
It will `git clone` the hooks from this repo into `.git/hooks` or similar. All
exported functions of this package are just for the users convenience.


## How can I make sure that my contributors are using the hooks?

They must follow the installation instructions in the README, i.e. run 
```{r}
remotes::install_github("lorenzwalthert/precommit")
precommit::install_precommit()
precommit::use_precommit()
```

The last call can be omitted by users who have [automatically enabled 
pre-commit hooks](https://pre-commit.com/#automatically-enabling-pre-commit-on-repositories).

To enforce all hooks pass, you can follow the [advice on how to use pre-commit in a CI/CD setup](https://pre-commit.com/#usage-in-continuous-integration).

## What if not all people who are committing to this repo want to use the hooks?

This is not a problem, git will only run the hooks in a local repo after
`precommit::use_precommit()` has been run successfully from within this local repo on your machine.
You can also uninstall anytime with `precommit::uninstall_precommit()`. Anyone
who does not want to use the hooks simply should not run
`precommit::use_precommit()`. You can also temporarily disable hooks as 
described [here](https://pre-commit.com/#temporarily-disabling-hooks).


## How does one create a new pre-commit hook?

How to contribute new hooks is explained in [`CONTRIBUTING.md`](https://github.com/lorenzwalthert/precommit/blob/master/CONTRIBUTING.md).

## Why are my hooks running slowly?

Some of the hooks depend on R functions that run slowly, e.g. the `style-files` 
hook depends on `styler`, which is not very fast. Also, note that for each hook,
R is started. Make sure your startup process is quick for non-interactive use, 
e.g. your `.Rprofile` and related files do not have a long execution time. E.g. 
wrap code in `if (interactive()) ` if possible, so it won't run when pre-commit 
hooks are ran. Find out more about the startup process with `?startup`. We'd 
like to support the `--vanilla` option but it's 
[not easily possible we believe](https://github.com/lorenzwalthert/precommit/issues/165).

*** 

There is more. Check out the documentation of the [pre-commit](https://pre-commit.com) framework.
