---
title: "Manual installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{manual-installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Installation 

Apart from `precommit::use_precommit()`, you can install pre-commit manually as
described in the [official documentation](https://pre-commit.com/#intro) (do not
close the log of this process, you might need it later!). Then, you must make
sure R finds the executable. To check that, start a new R session after the
installation process of pre-commit succeeded and run this command:

```{r, eval = FALSE}
precommit::path_pre_commit_exec()
#> "/usr/local/bin/pre-commit"
```
If it shows a path (like above), you are all set. Skip the rest of this
paragraph and go to *Configuration*. If not, you must manually 
link the executable.Set the R option
`precommit.executable` to the path where the pre-commit executable lives.
Note that the PATH variable that was used in the above command to find the 
executable might be different in your R session from what you
have in your shell, so you can try your luck there:

```bash
$where pre-commit
# should give /bin/pre-commit or similar
```

If this returns a path, set the R option to it, preferrably in your `.Rprofile`
to make sure it is set in a new R session too.

```{r, eval = FALSE}
options(precommit.executable = "/path/to/pre-commit")
```

Now, proceed in the below section on *Configuration*. If `$where pre-commit` did
not give you a path either, you need to go back to the installation method you
used and figure out where the executable got placed. This depends on the
installation method and oerating system, so it's hard for us to give general
advice. You may use a Google search like *where does pip install executable on
Ubuntu?* to find the path to the executable.

We are working towards making `precommit::path_pre_commit_exec()` finding
executables automatically with any operating system and installation method, so
please open up an [issue](https://github.com/lorenzwalthert/precommit/issues) on
GitHub and tell us your installation method, platform and the location of the
executable so we can adapt the search heuristic of
`precommit::path_pre_commit_exec()`.

If you can't get it righ, you can still install via conda, where linking the 
executable is not required.


## Configuration

If you've installed pre-commit, you can add it to a specific project by adding a
`.pre-commit-config.yaml` file that has a structure like this:

```{r,results='asis'}
tags <- names(git2r::tags("."))
latest_tag <- paste0("v", max(as.numeric_version(gsub("^v", "", tags))))
  
  
string <- paste0("``` yaml
repos:
-   repo: https://github.com/lorenzwalthert/precommit
    rev: ", latest_tag , "
    hooks: 
    -   id: style-files
    -   id: parsable-R
    -   id: no-browser-statement
    -   id: readme-rmd-rendered
    # -   id: roxygenize            # for repos that are a package.
    # -   id: use-tidy-description  # for repos that are a package.
    # -   id: deps-in-desc          # for repos that are a package.
    -   id: lintr
```")
cat(string)

```

The `repo` key points to the place where the source code of the hooks are
stored, not the repo you want to apply them to.

Some hooks also take arguments, see `vignette("hook-arguments")`.

If you want to see the file `.pre-commit-config.yaml` in RStudio, you have to
enable "Show Hidden Files" in the *Files* Pane of RStudio under *More*.

Next, run `pre-commit install` in your repo and you are done.
